Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-01-09 paper/paper.tex
fix histogram

For that, \tname{} parses the message looking for options associated
with the data. As described in Section~\ref{sec:rules-and-packets},
there is a correspondence between packet fields and rule
options. \tname{} leverages that mapping to infer options. The
examples of sections~\ref{sec:dos} and~\ref{sec:active-recon} show
cases where \tname{} infers options with this mapping. Unfortunately,
inferring options at the field granularity level is insufficient to
capture \CodeIn{content} options as the described on
Section~\ref{sec:content-example}. These options refer to parts of the
payload of the message and cannot be extracted directly from the
fields. Handling this case is critically important as \Fix{xx}\% of
the rules contain \CodeIn{content} options.

\pgfplotsset{width=6cm,compat=1.8}
\pgfplotsset{every tick label/.append style={font=\tiny}}
\begin{figure}[t!]
  \centering
  \scalebox{1.2}{  
    \begin{tikzpicture}
      \begin{axis}[
          bar width=6pt,
          scaled ticks=false,
          tick label style={/pgf/number format/fixed},
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          symbolic x coords={1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
          xtick=data,
          nodes near coords,
          every node near coord/.append style={font=\tiny},
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(1,10222) (2,4956) (3,5589)
          (4,1651) (5,1457) (6, 729) (7, 471) (8, 313) (9, 137) (10, 126)};
      \end{axis}
    \end{tikzpicture}
  }
  \caption{\label{fig:distribution-contents}Histogram of number of
    \CodeIn{content} options per rule.}
\end{figure}

Figure~\ref{fig:distribution-contents} shows the histogram of the
number of these options per rule obtained from a public
ruleset~\cite{emerging-threats-open} containing 27.8K Suricata
rules. We limited the number of content options per rule to 10 for
space. Rules without \CodeIn{content} options is an exception
(\Fix{xx}\%) and half of the rules contain at least two
\CodeIn{content} options.  To find the values associated with these
options, \tname{} splits the payload of the malicious message in
tokens using natural language delimiters, such as $\backslash$t,
$\backslash$n, $\backslash$r, and spaces. We observed, by inspecting
existing rules, that \CodeIn{content} options typically refer to
sequence of characters in the payload of messages separated by these
delimiters. This tokenization step can generate dozens of options.
\tname{} produced a total of \Fix{xx} content options for the example
Section~\ref{sec:content-example}.

@@

\begin{figure}[H]
  \centering
  \scalebox{0.85}{
    \begin{tikzpicture}
      \begin{axis}[
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          %          ylabel={\#rules},
          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
          xtick=data,
          xticklabel style={rotate=45},          
          nodes near coords,
          every node near coord/.append style={font=\scriptsize},        
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(0,2134) (1,9164) (2,5103) (3,5841) (4,1987) (5,1622) (6,805) (7, 535) (8, 340) (9,154) (10, 53)};
      \end{axis}
    \end{tikzpicture}
  }
  \vspace{-2ex}
  \caption{\label{fig:distribution-rules-protocol}Number of rules per number of contents.}
\end{figure}

>>>>

For that, \tname{} parses the message looking for options associated
with the data. As described in Section~\ref{sec:rules-and-packets},
there is a correspondence between packet fields and rule
options. \tname{} leverages that mapping to infer options. The
examples of sections~\ref{sec:dos} and~\ref{sec:active-recon} show
cases where \tname{} infers options with this mapping. Unfortunately,
inferring options at the field granularity level is insufficient to
capture \CodeIn{content} options as the described on
Section~\ref{sec:content-example}. These options refer to parts of the
payload of the message and cannot be extracted directly from the
fields. Handling this case is critically important as 93\% of
the rules contain \CodeIn{content} options.

\pgfplotsset{width=6cm,compat=1.8}
\pgfplotsset{every tick label/.append style={font=\tiny}}
\begin{figure}[t!]
  \centering
  \scalebox{1.2}{  
    \begin{tikzpicture}
      \begin{axis}[
          bar width=6pt,
          scaled ticks=false,
          tick label style={/pgf/number format/fixed},
          ybar stacked,
          enlargelimits=0.15,
          legend style={at={(0.5,-0.15)},
            anchor=north,legend columns=-1},
          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
          xtick=data,
          nodes near coords,
          every node near coord/.append style={font=\tiny},
          nodes near coords align={vertical},
        ]
        \addplot coordinates {(0,2134) (1,9164) (2,5103) (3,5841) (4,1987) (5,1622) (6,805) (7, 535) (8, 340) (9,154) (10, 53)};
      \end{axis}
    \end{tikzpicture}
  }
  \caption{\label{fig:distribution-contents}Histogram of number of
    \CodeIn{content} options per rule.}
\end{figure}

Figure~\ref{fig:distribution-contents} shows the histogram of the
number of these options per rule obtained from a public
ruleset~\cite{emerging-threats-open} containing 27.8K Suricata
rules. We limited the number of content options per rule to 10 for
space. Rules without \CodeIn{content} options is an exception
(7\%) and half of the rules contain at least two
\CodeIn{content} options.  To find the values associated with these
options, \tname{} splits the payload of the malicious message in
tokens using natural language delimiters, such as $\backslash$t,
$\backslash$n, $\backslash$r, and spaces. We observed, by inspecting
existing rules, that \CodeIn{content} options typically refer to
sequence of characters in the payload of messages separated by these
delimiters. This tokenization step can generate dozens of options.
\tname{} produced a total of \Fix{xx} content options for the example
Section~\ref{sec:content-example}.

@@

%%\begin{figure}[H]
%%  \centering
%%  \scalebox{1.2}{
%%    \begin{tikzpicture}
%%      \begin{axis}[
%%          ybar stacked,
%%          enlargelimits=0.15,
%%          legend style={at={(0.5,-0.15)},
%%            anchor=north,legend columns=-1},
%%          %          ylabel={\#rules},
%%          symbolic x coords={0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10},
%%          xtick=data,
%%          xticklabel style={rotate=45},          
%%          nodes near coords,
%%          every node near coord/.append style={font=\scriptsize},        
%%          nodes near coords align={vertical},
%%        ]
%%        \addplot coordinates {(0,2134) (1,9164) (2,5103) (3,5841) (4,1987) (5,1622) (6,805) (7, 535) (8, 340) (9,154) (10, 53)};
%%      \end{axis}
%%    \end{tikzpicture}
%%  }
%%  \vspace{-2ex}
%%  \caption{\label{fig:distribution-rules-protocol}Number of rules per number of contents.}
%%\end{figure}
